---
title: "Google Analytics Capstone Case Study 2"
author: "Shreshth Vashisht"
date: "2023-06-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Google Capstone

I used 4 datasets downloaded from the Kaggle dataset [Fitbit Fitness Tracker](https://www.kaggle.com/datasets/arashnic/fitbit) out of 18 datasets
They are 
* dailyActivity_merged
* sleepDay_merged
* dailyCalories_merged
* dailySteps_merged

These datasets were cleaned using Excel, where an additional column was added which was of correct date format


```{r importing dataset}
library(readr)
daily_act <- read_csv("C:/Users/Hp/Downloads/archive (1)/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv", col_types = cols(ActivityDateCorrected = col_date(format = "%d-%m-%Y")))
sleep <- read_csv("C:/Users/Hp/Downloads/archive (1)/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv", col_types = cols(SleepDay = col_character()))
cal <- read_csv("C:/Users/Hp/Downloads/archive (1)/Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv", 
col_types = cols(ActivityDayCorrected = col_date(format = "%d-%m-%Y")))
steps <- read_csv("C:/Users/Hp/Downloads/archive (1)/Fitabase Data 4.12.16-5.12.16/dailySteps_merged.csv", col_types = cols(ActivityDayCorrected = col_date(format = "%d-%m-%Y")))

```

Average Steps, Distance, Calories Burnt and Sedentry Mintues were calculated from the daily_act dataset

```{r calculating_average}
d_act_avg<- daily_act %>% group_by(Id) %>% drop_na() %>% summarise(avg_total_steps=mean(TotalSteps), avg_distance=mean(TotalDistance), avg_calories_burnt=mean(Calories), avg_sedentary=mean(SedentaryMinutes))
print(paste("average total steps per person is",mean(d_act_avg$avg_total_steps)))
print(paste("average distance travelled per person is", mean(d_act_avg$avg_distance)))
print(paste("average calories burnt per person is",mean(d_act_avg$avg_calories_burnt)))
print(paste("average Sedentary Minutes per person is",mean(d_act_avg$avg_sedentary)))
```
Finding Relationship
```{r correlations}
a<-d_act_avg %>% summarise(cor(avg_total_steps,avg_calories_burnt))
b<-d_act_avg %>% summarise(cor(avg_distance,avg_calories_burnt))
d<-d_act_avg %>% summarise(cor(avg_calories_burnt,avg_sedentary))
print(paste("Relationship between Steps and Calories Burnt:",a))
ifelse(a>0.7,"High Corelation","Low Corelation")
print(paste("Relationship between Distance and Calories Burnt:",b))
ifelse(b>0.7,"High Corelation","Low Corelation")
print(paste("Relationship between Sedentary Minutes and Calories Burnt:",d))
ifelse((-0.7>d|d>0.7),"High Corelation","Low Corelation")
```
Now we will view the Plots

```{r Plot1}
ggplot(d_act_avg)+ geom_point(mapping = aes(x=avg_total_steps,y=avg_calories_burnt))+geom_smooth(mapping = aes(x=avg_total_steps,y=avg_calories_burnt))
```

```{r Plots2}
ggplot(d_act_avg)+geom_point(mapping = aes(x=avg_distance,y=avg_calories_burnt)) +geom_smooth(mapping = aes(x=avg_distance,y=avg_calories_burnt))
```

```{r Plots3}
ggplot(d_act_avg)+geom_point(mapping = aes(x=avg_distance,y=avg_total_steps)) +geom_smooth(mapping = aes(x=avg_distance,y=avg_total_steps))
```
```{r Plots4}
ggplot(d_act_avg)+ geom_point(mapping = aes(x=avg_sedentary,y=avg_calories_burnt))+geom_smooth(mapping = aes(x=avg_sedentary,y=avg_calories_burnt))
```
Now we'll view the plots from sleep dataset

```{r sleep}
act_date_avg<- daily_act %>% group_by(ActivityDateCorrected) %>% drop_na() %>% summarise(avg_total_steps=mean(TotalSteps), avg_distance=mean(TotalDistance), avg_calories_burnt=mean(Calories), avg_sedentary=mean(SedentaryMinutes))
```

Now we will find relationship
```{r correlations2}
a<-act_date_avg %>% summarise(cor(avg_total_steps,avg_calories_burnt))
b<-act_date_avg %>% summarise(cor(avg_distance,avg_calories_burnt))
d<-act_date_avg %>% summarise(cor(avg_calories_burnt,avg_sedentary))
print(paste("Relationship between Steps and Calories Burnt:",a))
ifelse(a>0.7,"High Corelation","Low Corelation")
print(paste("Relationship between Distance and Calories Burnt:",b))
ifelse(b>0.7,"High Corelation","Low Corelation")
print(paste("Relationship between Sedentary Minutes and Calories Burnt:",d))
ifelse((-0.7>d|d>0.7),"High Corelation","Low Corelation")
```
This is an interesting observation that per user there is low correlation between distance and calories burnt and between steps and calories burnt and this needs to be further investigated


Now we will view PLots

```{r Plots5}
ggplot(act_date_avg)+ geom_point(mapping = aes(x=avg_total_steps,y=avg_calories_burnt))+geom_smooth(mapping = aes(x=avg_total_steps,y=avg_calories_burnt))
```
```{r Plots6}
ggplot(act_date_avg)+geom_point(mapping = aes(x=avg_distance,y=avg_calories_burnt)) +geom_smooth(mapping = aes(x=avg_distance,y=avg_calories_burnt))
```
```{r Plots7}
ggplot(act_date_avg)+geom_point(mapping = aes(x=avg_distance,y=avg_total_steps)) +geom_smooth(mapping = aes(x=avg_distance,y=avg_total_steps))
```


```{r Plots8}
ggplot(act_date_avg)+ geom_point(mapping = aes(x=avg_sedentary,y=avg_calories_burnt))+geom_smooth(mapping = aes(x=avg_sedentary,y=avg_calories_burnt))
```

We need to investigate why in per person dataset there is negative correlation between sedentary and calories burnt but positive in date wise dataset
